---
title: "Prediction Assignment"
output: html_document
---
## Synopsis
One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. This project will use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants.  The goal is to use that data to predict the manner in which the participants did the exercise:

    Class A: exactly according to the specification
    Class B: throwing the elbows to the front
    Class C: lifting the dumbbell only halfway
    Class D: lowering the dumbbell only halfway
    Class E: throwing the hips to the front

For more information, see http://groupware.les.inf.puc-rio.br/har

## Prepare Data
### Read data and see number of rows and columns
```{r}
alltrain = read.csv("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv")
testing = read.csv("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv")

dim(alltrain)
dim(testing)

```

### Slice training into training and validation
```{r}
library(caret)
set.seed(1342)
inTrain <- createDataPartition(y=alltrain$classe,p=0.7, list=FALSE)
training <- alltrain[inTrain,]; 
validation <- alltrain[-inTrain,]

```

### Take a quick look at training data
```{r eval=FALSE}
head(training)
```
The first seven columns are identifiers such as names and time stamps, so I will remove them before training.  I will also remove columns that don't have numeric values for all rows.
```{r}
library(caret)
classe <- training$classe
training <- training[,-seq(1:7)]
training <- training[, colSums(is.na(training)) == FALSE]
training <- training[, sapply(training, is.numeric)]
training$classe <- classe
dim(training)
summary(training$classe)
```
```{r eval=FALSE}
head(training)
```

## Fit a Model
Just as in the original paper, I will use Random Forest algorithm
```{r}
library(ggplot2)
library(rattle)
set.seed(1837)
modFit <- train(classe ~ ., method="rf", data=training, ntree = 50)
modFit
# fancyRpartPlot(modFit$finalModel)
```


## Model Performance
### With Validation set
The accuracy is estimated at 99.17%
```{r}
valdPred <- predict(modFit, validation)
confusionMatrix(valdPred, validation$classe)
```

### With Testing set
When this model was used against the testing data, it scored 20 out of 20 correct.
```{r}
predict(modFit,newdata=testing)
```

